### 2015-10-20 p01-ch01-2012/langs.tex

\documentclass{article}
\usepackage[top=0.01in,bottom=0.01in,left=0.01in,right=0.01in,paperwidth=5.7in,paperheight=2.3in]{geometry}

\usepackage{alltt}
\usepackage{array}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{url}
\usepackage[table]{xcolor}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand{\thickhline}{\noalign{\hrule height 1.0pt}}

\begin{document} 
\pagestyle{empty}
\definecolor{lgray}{rgb}{0.85,0.85,0.85}
\rowcolors{2}{lgray}{white}
\begin{center} 
\begin{tabular}{m{0.5in} m{1.2in} m{1in} m{1in} m{1in}}
  & Java & C++ & Python & C\# \\
\thickhline 
Actors &  Scala, Erlang, SALSA, Axum, Ptolemy, Kilim, ActorFoundry, Jetlang,
JavAct, AJ & Theron, Act++ & Stackless, V3.2+ & Retlang \\

STM & Multiverse, Clojure, DSTM2 (Fortress), Deuce, JVSTM & TL2 (Sun), Intel
C++ STM & Durus & SXM \\

Data Flow & Java 7+ JSR 166y Fork/Join, Hadoop & OpenMP, TBoost.STM, Hadoop Streaming &
Disco (Nokia) & Dryad, Hadoop Streaming \\
\end{tabular}
\end{center}
\end{document}