### 2015-10-20 p01-ch01-2012/paper.tex

% -*- root: ../thesis.tex -*-

\chapter[Application-Level Scheduling]{Programming and Deployment of Active Objects with Application-Level Scheduling%
\footnote{This work is partially supported by the EU FP7-231620 project: HATS.}
}
% 
\label{ch:p01:ch01}
% 
\chapterauthor{Behrooz Nobakht, Frank S. de Boer, Mohammad Mahdi Jaghoori,  Rudolph Schlatte}

% --- Author Metadata here ---
% \conferenceinfo{SAC'12}{March 25-29, 2012, Riva del Garda, Italy.}
% \CopyrightYear{2011}
% \crdata{978-1-4503-0857-1/12/03}
% --- End of Author Metadata ---

% \title{Programming and Deployment of Active Objects with Application-Level Scheduling%
% \titlenote{This work is partially supported by the EU FP7-231620 project: HATS.}
% }

% \numberofauthors{2}  %set to two to enable longer lines for names and

% \author{
% \alignauthor Behrooz Nobakht\\
% 	\affaddr{Leiden Institute of Advanced Computing}\\
% 	\affaddr{Universiteit Leiden, Leiden, The Netherlands}\\
% 	\email{bnobakht@liacs.nl}
% \alignauthor Frank S. de Boer\\
% 	\affaddr{Centrum Wiskunde and Informatica}\\
% 	\affaddr{Amsterdam, The Netherlands}\\
% 	\email{frb@cwi.nl}
% \and
% \alignauthor Mohammad Mahdi Jaghoori\\
% 	\affaddr{Centrum Wiskunde and Informatica}\\
% 	\affaddr{Amsterdam, The Netherlands}\\
% 	\email{jaghoori@cwi.nl}
% \alignauthor Rudolf Schlatte\\
% 	\affaddr{University of Oslo}\\
% 	\affaddr{Oslo, Norway}\\
% 	\email{rudi@ifi.uio.no}
% }

\section*{Abstract}
% Active objects provide a natural means for programming distributed systems.
We extend and implement a modeling language based on concurrent active objects with application-level
scheduling policies.
The language allows a programmer to assign priorities at the application level, for example,
to method definitions and method invocations, and assign corresponding  policies to the individual active objects
for scheduling the messages.
Thus, we   leverage scheduling and performance related issues,
which are becoming increasingly important in multi-core and cloud  applications, from the underlying operating system to the application level.
We describe a tool-set to transform models of active objects extended with application-level
scheduling policies  into Java.
This tool-set allows a direct  use of Java class libraries; thus,  we
obtain  a full-fledged programming language
based on active objects which allows for high-level control of deployment related issues.

% \category{D.1.3}{Programming Techniques}{Concurrent Programming}
% \category{D.2.3}{Software Engineering}{Coding Tools and Techniques}
% \category{D.3.0}{Programming Languages}{General}
% \terms{Languages, Design}
% \keywords{Application-level scheduling, ~Priority scheduling, ~Creol, Java, Actor model, Concurrent active objects}

\section{Introduction} \label{sec:introduction}
One of the major challenges in the design of programming
languages is to provide high-level support for  multi-core and cloud  applications which are becoming increasingly important.
Both multi-core and cloud  applications require an explicit and precise treatment of non-functional properties, e.g., resource requirements.
On the cloud, services execute in the
context of virtual resources, and the amount of resources actually available to a service is subject to change.
Multi-core applications require techniques 
to help the programmer optimally use potentially many cores. At
the operating system level, resource management is greatly affected by scheduling which
is largely beyond the control of most existing high-level programming languages.
Therefore, for optimal use of the available resources, we cannot avoid leveraging
scheduling and performance related issues from the underlying operating system
to the application level. However, the very nature of high-level languages is to
provide suitable abstractions that hide
implementation details from the programmer. The main challenge in designing programming languages
for  multi-core and cloud  applications
is to find a balance between these two conflicting requirements.

We investigate in this paper how concurrent active objects in a high-level object-oriented
language can be used for high-level scheduling of resources. We use the notion of
concurrent objects in Creol \cite{creol:broch_owe,mpd:andrews}. A
concurrent object in Creol has control over one processor; i.e. it has a
single thread of execution that is controlled by the object itself.
Creol processes never leave the enclosing object; method invocations
result in a new process inside the target object.  Thus, a concurrent
object provides a natural basis for a deployment scheme where each
object virtually possesses one processor. Creol further provides high-level
mechanisms for synchronization of the method invocations in an object;
however, the scheduling of the method invocations are left
unspecified. Therefore, for the deployment of concurrent objects in
Creol, we must, in the very first place, resolve the
basic scheduling issue; i.e. which \textit{method} in which
\textit{object} to select for execution.
We show how to introduce priority-based scheduling of the messages of the individual objects
at the application-level itself.

In this paper we also propose a 
% pluggable and extensible 
tool
architecture to deploy Creol applications. To prototype the tool
architecture, we choose Java as it provides low-level concurrency features, i.e., threads, futures, etc.,
required for multi-core deployment of object-oriented applications. The tool
architecture prototype transforms Creol's constructs for concurrency to their
equivalent Java constructs available in the
{\jtt{java.util.\-concurrent}} package. As such, Creol provides
a high-level structured programming discipline based on active objects on top of Java.
Every active object in Creol is transformed to an object in Java that uses a priority manager
and scheduler to respond to the incoming messages from other objects. Besides,
through this transformation, we allow the programmer to seamlessly use, in the
original Creol program, Java's standard library including all the data types.
Thus, our approach converts Creol from  a modeling language to a full-fledged
 ``programming'' language.

Section \ref{sec:creol} first provides an overview of the Creol language with application-level scheduling. 
In Section \ref{sec:compiler}, we elaborate on the design  of the tool-set and the
prototype. The use of the tool-set is exemplified by a case study in Section \ref{sec:caseStudy}.
Section \ref{sec:relwork} summarizes the related work. 
Finally, we conclude in Section \ref{sec:conclusion}.

%\section{Preliminaries: Creol} \label{sec:creol}
%\input{Creol}

\lstset{language=Creol,escapeinside={(*}{*)}}

\section{Application-Level Scheduling}
\label{sec:creol}

Creol \cite{creol:broch_owe} is a full-fledged object-oriented language with formal semantics for modeling and analysis of systems of concurrent objects. 
Some Creol features include interface and class inheritance and being strongly typed such that safety of dynamic class upgrades can be statically ensured \cite{yu06fmoods}.
In this section, we explain the concurrency model of Creol using a toy example: an exclusive resource, i.e., a
resource that can be exclusively allocated to one object at a time, behaving as a mutual exclusion token. 
Further, we extend Creol with priority-based application-level scheduling.

The state of an object in Creol is initialized using the \lstinline{init} method.
Each object then starts its active behavior by executing its
\lstinline{run} method if defined. % (e.g., in class \lstinline{User}). 
When receiving a method call, a new process is created to execute the method. 
Creol promotes {\em cooperative} non-preemptive scheduling for each active object. 
It means that a method runs to completion unless it explicitly releases the processor. 
As a result, there is no race condition between different processes accessing object variables. 
%Methods can have processor release points which define interleaving points explicitly.
Release points can be conditional, e.g.,   \lstinline{await }$\sim$\lstinline{taken}. 
If the guard at a release point evaluates to true, the
process keeps the control, otherwise, it releases the processor and becomes
disabled as long as the guard is not true. Whenever the processor is free, an
enabled process is {\em nondeterministically} selected for execution, i.e.,
scheduling is left unspecified in standard Creol in favor of more abstract modeling.

%\Mahdi{Do we want to talk about with clauses?}

To explain extending Creol with priority specification and scheduling, we take a {\em client/server} perspective.
Each {caller} object is viewed as a client for the {callee} object who behaves as a server. 
We define priorities at the level of language constructs like method invocation or definition rather than low-level concepts like processes.

\begin{lstlisting}[float=t, label=lst:exres, caption=Exclusive Resource in Creol %, multicols=2
]
interface Resource begin
	(* \label{lst:priority_range} *)(*
	%with User  \label{lst:with} 
	*)
		op request()
		op release()
end

class ExclusiveResource implements Resource begin
	var taken := false;
	(* \label{lst:priority_class} *)
	op request () == 
		await (*$\sim$*)taken; 			(* \label{lst:await_taken} *)
		taken := true;
	op release () ==
		taken := false
end
\end{lstlisting}

On the server side, an interface may introduce a \textit{priority
range} that is available to all clients. For instance, in Line
\ref{lst:priority_range} of {\footnotesize\texttt{Resource}} in Listing
\ref{lst:exres}, we can define a priority range:

\begin{lstlisting}[frame=none, numbers=none]
priority range 0..9
\end{lstlisting}

On the client side, method calls may be given priorities within the range specified in the server interface. For example, calling the \ftt{request} method of the \ftt{mutex} object:

\begin{lstlisting}[frame=none, numbers=none]
mutex ! request() priority(7);
\end{lstlisting}

Scheduling \textit{only} on the basis of client-side priority requirements is too restrictive.
For example, if there are many \ftt{request} messages with high priorities and a low priority \ftt{release}, the server might block as it would fail to schedule the \ftt{release}.
In this particular example, we can solve this problem by allowing the servers to
prioritize their methods.
This involves a declaration of a priority range generally depending on the structure of the class. 
In our example, assuming a range 0..2 added in Line \ref{lst:priority_class}, this requires changing the method signatures in the {\ftt{Exclusive\-Resource}} class:

\begin{lstlisting}[frame=none, numbers=none]
op request() priority(2) == ...
op release() priority(0) == ...
\end{lstlisting}

This gives \ftt{release}  a higher priority over \ftt{request}, because by default, the smaller values indicate higher priorities.
Furthermore, the server may also introduce priorities on certain 
characteristics of a method invocation such as the kind of ``release statement''
being executed. 
For example, a process waiting at Line \ref{lst:await_taken} could be given a higher priority over new requests by assigning a smaller value:

\begin{lstlisting}[frame=none, numbers=none]
await (*$\sim$*)taken priority(1);
\end{lstlisting}

The priority can be specified in general as an expression; we used here
only constant values.
Evaluation of this expression at runtime should be within the given
range.
If no priority is specified for a method invocation or definition, a
default priority value will be assigned.

% 
% This is an example of {\em method level} priority specification, which is specified inside a class together with the definition of a method. 
% Method definitions without an explicit priority are given a default priority value. 
% %As discussed in Section \ref{sec:conclusion}, 
% Other possible levels of priority include method invocation level, specified by the caller, or interface level priorities. 
% %

We discussed different levels of application-level priorities. 
Note that now each method invocation in
the queue of the object involves a tuple of priorities.
We define a general function $\delta$ as an ``abstract priority manager'':
%
\[ \delta : P_1 \times P_2 \times P_3 \times \ldots \times P_n \longrightarrow  P \]

The function $\delta$ maps the different levels of priority in the object
($\{P_1,\ldots,P_n\}$) to a single priority value in $P$ that is internally used by the
object to order all the messages that are queued for execution. Each
method in the object may have a $\delta$ assigned to it. In an extended
version of $\delta$, it can also involve the internal state of the object,
e.g., the fields of the object. 
In this case, we have dynamic priorities.

For example, in {\ftt{ExclusiveResource}}, we have two different levels of priorities, namely  the client-side and server-side priorities, which range over $P_1 = \{0, \ldots, 9\}$ and $P_2 = \{0, 1, 2\}$, respectively. 
So, we define $\delta: P_1 \times P_2 \rightarrow P$ as:
%
\[
\delta(p_1, p_2) = p_1 + p_2 \times |P_1|
\]

To see how it works, consider a \ftt{release} and a \ftt{request} message, both sent with the client side priority of 5. 
Considering the above method priorities, we have $\delta(5, request) = \delta(5,2) = 25$ and  $\delta(5, release) = \delta(5,0) = 5$.
It is obvious that the range of the final priority value is $P = \{0,\ldots,29\}$.

Note that the abstract priority manager in general does \textit{not} completely
fix the ``choice'' of which method invocation to execute. 
In our tool-set, we include an extensible library of predefined scheduling policies such as strong or weak fairness that further refine the application-specific multi-level priority scheduling. 
The policies provided by the library are available to the user to annotate the classes.
We may declare a scheduling policy for the \ftt{ExclusiveResource} class by adding at Line
\ref{lst:priority_class} in Listing \ref{lst:exres}: 
%
\begin{quote}
\lstinline{scheduling policy StronglyFairScheduler;} 
\end{quote}

% These scheduling policies will also be represented in terms of a corresponding
% priority range in the $\delta$ function.
% In summary, there will be, first, priorities that directly stem from the
% model and are statically specified by the client or the server and, second, 
% there will be priorities that dynamically change through time based on
% the runtime information of the application. In the latter case, 
A scheduling policy may
use runtime information to re-compute the dynamic priorities and ensure
properties such as fairness of the selected messages; for instance, it may
take advantage of the ``aging'' technique to avoid starvation. 
%Thus, $\delta$ will take into consideration all the different concerns when assigning a final priority value to a message.

%\Mahdi{Starvation?}

\section{Tool Architecture} \label{sec:compiler}

\begin{figure}[t]
\begin{center}
  %\vspace{-.1cm}
  \includegraphics[scale=0.36]{figs/manticore-concurrency-v4}
  \caption{\Crisp Architecture: Structural Overview}
  \label{fig:api}
\end{center}
\end{figure}

We have implemented a tool to translate Creol programs into Java programs for execution, called {\em \Crisp} (Creolized Interacting Scheduling Policies).
\Crisp provides a one-to-one mapping from Creol classes and methods to their equivalent Java constructs. 
In order to implement active objects in Creol, we use the \jtt{java.util.concurrent} package (see Figure \ref{fig:api}).
Each active object consists of an instance of a process store  and an execution manager to store and execute the method invocations.

Incoming messages to the active object are modeled as instances of \MethodInvocation, a subclass of \jtt{java.util.\-concurrent.FutureTask} that  wraps around the original method call.
Therefore the caller can later get the result of the call.
Additionally, \MethodInvocation encapsulates information such as priorities assigned to the message. 
% Asynchronous messages are stored in the process store as instances of \MethodInvocation. The result of the asynchronous method call can be fetched through the method invocation that encapsulates a future value.
% To support Creol's processor release statements, method invocations provide the
% means to be stopped and resumed.

The \ftt{ProcessStore} itself uses an implementation of the \ftt{BlockingQueue} interface in \jtt{java.util.concurrent} package. Implementations of \ftt{BlockingQueue} are thread-safe, i.e.,  all methods in this interface operate atomically using internal locks encapsulating the implementation details from the user.

The \ftt{ExecutionManager} component is responsible for selecting and executing a pending method invocation.
It makes sure that only one method runs at a time, and takes care of processor release points.
%It is implemented using \jtt{java.util.concurrent} in order to provide deployment on multi-core  architectures. We now discuss how in overall an active object operates on a multi-core architecture.

In the following, we explain how the active object behaves in different scenarios, from a ``client/server'' perspective.  
% for the sake of explanation; each active object
% essentially can be seen as a server, in which case all the other objects act as its clients that
% request a method to be executed.

% \subsection{The Problem}
% \input{problem}

\subsection{A New Method Invocation}
\label{sec:add}

\begin{figure}[t]
%\begin{center}
\includegraphics[scale=0.36]{figs/seq-diag-1}
~\\[-3.5cm]ADD:\\[2.5cm]
  \caption{Adding the new MethodInvocations are performed on the Client side.}
  \label{fig:add}
%\end{center}
\end{figure}

A method call needs to be sent from the client to the server in an asynchronous way. 
To implement this in Java, the client first constructs an instance of
\MethodInvocation that wraps around the original method call for the server.
Then, 
%the client needs to directly fetch the \ftt{ProcessStore} of the
%server and \ftt{add}s the instance of the method invocation to the server.
%
%Active objects consist of a \texttt{\textbf{ProcessStore}} which stores
% method invocations as described above and a \texttt{\textbf{ExecutionManager}}
% which controls the actual execution of method invocations.
% When an instance of \texttt{\textbf{MethodInvocation}} is created by a client, 
there are two implementation options how to add it to the server's process store:
\begin{enumerate}
  \item The client calls a method on the server to store the instance.
  \item The client directly adds the method invocation into the process
store of the server. 
\end{enumerate}

In option $1$, the server may be busy doing something else. Therefore,
in this case the client must wait until the server is free,  which is
against the asynchronous nature of communication. In Option $2$, the
Java implementation of each active object exposes its process store as
an immutable public property. Thus, the actual code for adding the new
method invocation instance is run in the execution thread of the
client. 
We adopt the second approach as depicted in Figure \ref{fig:add}.
At any time, there can be concurrent clients that are
storing instances of \ftt{MethodInvocation} into the server's process store, but since
the process store implementation encapsulates the mechanisms for concurrency and
data safety, the clients have no concern on data synchronization and concurrency
issues such as mutual exclusion.
The method or policy used to store the method invocation in the process store of
the server is totally up to the server's process store implementation details.
%Storing a method invocation in the process store of the server takes place in
%the client side thread of execution. Listing \ref{lst:jfib_method_fibonacci}
%presents a sample generated Java code for an asynchronous method call in Listing
%\ref{lst:cfib}.

\subsection{Scheduling the Next Method Invocation}

On the server side of the story,  an active object 
repeatedly fetches an instance of method invocation from its process
store for execution (cf. Fig \ref{fig:take}).
The process store uses its instance of
\ftt{SchedulingManager} to choose one of the method invocations. 
\Crisp has some predefined scheduling policies that can be used as scheduling managers;
nevertheless, new scheduling policies can be easily developed
and customized based on the requirements by the user. 
%Listing \ref{lst:aoloop} presents the main loop for the active object.

\begin{figure}[t]
~\\TAKE:\\
%\begin{center}
%\vspace{-1.3cm}
  \includegraphics[scale=0.3]{figs/seq-diag-2}
\vspace{-.7cm}
  \caption{An active object selects a method invocation based on its local scheduling policy. After a method finishes execution, the whole scenario is repeated.}
  \label{fig:take}
%\end{center}
\end{figure}

\ftt{SchedulingManager} is an interface the implementations of which
introduce a function to \emph{select} a method invocation based on different
possible criteria (such as time or data) that is either predefined or customized
by the user. The scheduler manager is a component used by process store when
asked to remove and provide an instance of method invocation to be executed.
Thus, the implementation of the scheduling manager is responsible how to choose
one method invocation out of the ones currently stored in the process store of
the active object. Different flavors of the scheduling manager may include
time-based, data-centric, or a mixture.
%\Mahdi{What are concrete examples of scheduling managers?} 
%\Rudi{There is a paper submitted to FMCO post-proceedings about user-defined
%scheduling policies; unfortunately there is no reference to cite yet}

Every method invocation may carry different levels of priority information,
e.g., a server side priority assigned to the method or a
client side priority. 
The \ftt{PriorityManager} 
%is an interface the implementations of which are supposed to 
provides a function to determine and
resolve a final priority value in case there are different levels of priorities
specified for a method invocation. 
% This feature is used, if specified, by the
% process store to compute the ``resolved priority'' of the method invocation
% based on different levels of priorities specified and then add it to the
% storage. 
%An example of a user-defined priority manager is given in Section \ref{sec:caseStudy}.
Postponing the act of resolving priorities to this point rather than when inserting new processes to the store enables us to handle dynamic priorities.

\subsection{Executing a Method Invocation}

To handle processor release points, Creol processes should preserve
their state through the time of awaiting.
This is solved by assigning an instance of a Java thread to each method
invocation. 
An \ftt{ExecutionManager} instance, therefore, employs a ``thread pool''
for execution of its method invocations. 
To create threads, it makes use of the factory pattern:
\ftt{ThreadFactory} is an interface used by the execution manager to
initiate a new thread when new resources are required. 
We cache and reuse the threads so that we can control and tune the performance of resource allocation.

When a method invocation has to release the processor, its thread must
be suspended and, additionally, its continuation must be added to the
process store. To have the continuation, the thread used for the method
invocation should be preserved to hold the current state; otherwise the
thread may be taken away and the continuation is lost.
The original \ftt{wait} in Java does not provide a way to achieve this
requirement.
Therefore, we introduce \ftt{InterruptibleProcess} as an extension of
\jtt{java.lang.Thread} to preserve the relation.

As shown in Figure \ref{fig:execute}, the thread factory creates threads of type  {\ftt{InterruptibleProcess}}.
The execution manager thread blocks as soon as it starts the interruptible process which executes the associated  method invocation. 
If the method releases the processor before completion, it will be added back to the process store as explained in Section \ref{sec:add}.
When a suspended method invocation is resumed, the execution manager skips the creation of a new thread and  reuses the one that was assigned to the method invocation before.

\begin{figure}[t]
~\\EXECUTE:\\
% \begin{center}
% \vspace{-1.3cm}
 \includegraphics[scale=0.35]{figs/seq-diag-3}
\vspace{-.7cm}
\caption{A method invocation is executed in an interruptible process. The execution manager thread is blocked while the interruptible process is running.}
  \label{fig:execute}
% \end{center}
\end{figure}

% Our \ftt{ExecutionManager} is an implementation of the standard Java ~interface 
% \jtt{java.util.concurrent.Executor\-Service}. 
% Each active object holds a reference to one execution manager instance.
% Each \ftt{ActiveObject} implements \jtt{java.lang.\-Run\-na\-ble}. So, when
% it gets initialized, it adds \ftt{this} (itself) to
% its execution manager component. This procedure \textit{deploys} the active object
% onto the execution manager corresponding to the underlying multi-core platform.

% FRB: the above last  sentence is not very clear and seems more to hide than to clarify.
% Also the remainder of the section needs to be improved.
% I suggest to add some sequence diagrams.
% Question: when a method invocation executes an await what actually happens to its executing thread?
% Why do we need a thread-pool because only one method at a time is executed?

\subsection{Extension Points}

Besides the methods \ftt{add} and \ftt{take} for adding and removing method invocations, \ftt{ProcessStore} provides methods such as \ftt{preAdd} and \ftt{postAdd} along with \ftt{preTake} and \ftt{postTake} respectively to enable  further customization of the behavior before/after adding or taking a method invocation to/from the store. These extension points enable the customization of priority or scheduling management of the method invocations.

\Crisp provides two generic interfaces for priority specification and scheduling management: \ftt{PriorityManager} and \ftt{SchedulingManager} respectively. These two interface can be freely developed by the programmer to replace the generated code for priorities and scheduling of the messages. It will be the task of the programmer to configure the generated code to use the custom developed classes.

\section{Case Study} \label{sec:caseStudy}

In this section, we demonstrate the use of application-level scheduling and \Crisp with a more complicated example: we program the ``Sieve of Eratosthenes'' %\cite{Sieve_Eratosthenes_wiki}
to generate the prime numbers. 
To implement this algorithm, the 
% \ftt{Sieve} class generates the natural numbers and sends them to the \ftt{Prime} class instances to check whether they can divide these numbers: if a number is not divisible by all primes before that, then it is a new prime number.
% %Initialization of an object happens in its \ftt{init} method, while the \ftt{run} method specifies its active behavior.
% The 
\ftt{Sieve} is initialized by creating an instance of the \ftt{Prime} object representing the first prime number, i.e., \ftt{two}. The active behavior of \ftt{Sieve} consists of generating all natural numbers up to the given limit (100000 in our example) and passing them to the object \ftt{two}. 
A \ftt{Prime} object that cannot divide its input number passes it on to the next \ftt{Prime} object; if there is no next object, then the input number is the next prime and therefore a new object is created.

We parallelize this algorithm by creating active objects that run in parallel.
The numbers are passed asynchronously as a parameter to the \ftt{divide} message. 
Correctness of the parallel algorithm essentially depends on the numbers being processed in increasing order.
For example, if object \ftt{two} processes 9 before 3, it will erroneously treat 9 as a prime, because 3 is not there yet to eliminate 9. 
%Such a scenario is possible if the scheduling of the tasks in the \ftt{Prime} objects is not specified. 
%
To avoid erroneous behavior, we use the actual parameter $n$ in \ftt{divide} method to define its priority level, too (see line \ref{line:priority}). 
As a result, every invocation of this method generates a process with a priority equal to its parameter.
The default scheduling policy for objects always selects a process for execution that has the smallest priority value. 
This guarantees that the numbers sent to a \ftt{Prime} object are processed exactly in increasing order.

\lstset{language=Creol,escapeinside={(*}{*)}}  
\begin{lstlisting}[float=t, label=lst:prime-sieve-creol, caption=Prime Sieve in Creol]
interface IPrime begin
	op divide(n:Int)
end

class Sieve begin
	var n: Int, two: IPrime
	op init == two := new Prime(2); n := 3
	op run ==
		!two.divide(n);
		if n < 100000 then n := n + 1; !run() end
end

class Prime(p: Int) implements IPrime begin (*\label{lst:priority}*)
	var next: IPrime
	op divide(n: Int) priority (n) == (*\label{line:priority}*)
		if (n % p) /= 0 then 
			if next /= null then
				!next.divide(n)
			else
				next := new Prime(n)
		end end
end
\end{lstlisting}
% 
% class Test
% begin
% 	var g : Sieve
% 	op run == g := new Sieve
% end

We used two different setups to execute the prime sieve program and compare the results. In one setting, we ran the parallel prime sieve compiled by \Crisp; in the other, we executed a sequential program developed based on the same algorithm that uses a single thread of execution in JVM. We performed the experiments on a hardware with $2$ CPU's of each $2$GHz with a main memory of size $2$GB. We ran both programs for $max \in \{10000, 20000, 30000, 50000, 100000\}$.
% We compared two interesting properties of the executions: the thread stack and the number of threads required.

The first interesting observation was that \Crisp prime sieve utilizes all the CPU's on the underlying hardware as much as possible during  execution. 
This can be seen in Figure \ref{fig:multi-cpu-1} which shows the CPU usage. 
Both CPUs are fairly in use while running this program.
Figure \ref{fig:prime_sieve_cacoj} depicts the results of the monitoring of the parallel prime sieve in \Crisp using Visual VM tool. 
It depicts the number of threads generated for the program.
This shows that \Crisp can handle a massive number of concurrent tasks.

One interesting feature of \Crisp is that the execution of any program under \Crisp can constantly utilize the \emph{minimum} memory that can be allocated for each thread in JVM (thread stack). 
In JVM, the size of thread stack can be configured using \ftt{-Xss} option for every run. 
To demonstrate this feature of \Crisp, we collected the minimum stack size needed for every program run in Table \ref{tbl:cmp-thread-stack}. 
All \Crisp runs use the minimum thread stack size of \ftt{64k} that is possible for the JVM.
On the contrary, the stack size required for the sequential version of the sieve program increases by the number of primes detected. 
This is also expected because of the long chain of method calls in the sequential sieve.

% The results depict that the sequential version, through time, increasingly required memory allocation for the thread stack, however, \Crisp performs on a constant size of thread stack. 
% Table \ref{tbl:cmp-thread-stack} summarizes the thread stack allocation on different runs.

\begin{table}[b]\scriptsize
\begin{center}
\begin{tabular}{c c c c c c}
$max$ & $10000$ & $20000$ & $30000$ & $50000$ & $100000$ \\ \thickhline
Sequential & 64k & 72k & 96k & 160k & 190k \\ 
\Crisp & 64k & 64k & 64k & 64k & 64k \\ \thickhline
\end{tabular}
\caption{Thread stack allocated for different executions}
\label{tbl:cmp-thread-stack}
\end{center}
\end{table}

Having the constant thread stack size feature, \Crisp provides another interesting feature. It can handle huge number of thread generation if required. Table \ref{tbl:cmp-thread-num-gen} summarizes the thread generation data for parallel prime sieve in \Crisp. It shows scalability of \Crisp as $p$ rises for parallel prime sieve.
\begin{table}[b]\scriptsize
\begin{center}
\begin{tabular}{c c c}
$max$ & Live Peak & Total \\ \thickhline
$10000$ & 817 & 540591 \\
$20000$ & 1468 & 1854067 \\
$30000$ & 2204 & 4054814 \\
$50000$ & 3707 & 11852985 \\ \thickhline
\end{tabular}
\caption{Number of live threads and total threads created for different runs of parallel prime sieve}
\label{tbl:cmp-thread-num-gen}
\end{center}
\end{table}

As the results show the use of Java threads is costly; first, \Crisp
does not need much of the memory allocated to each thread and, second,
the context switch cost is higher for larger memory allocation. In line
with this, JVM uses a \emph{one-to-one} mapping from an
application-level Java thread to a native OS-level thread. 
In the current setting, the context switch of the threads are in
the OS level. When the context switch is taken to the application
level, we leverage the performance issue from the OS level to the
application level. We further discuss this in Section
\ref{sec:conclusion}.

% In future work, in order to support high numbers of concurrent tasks,
% \Crisp will map a set of Creol messages to one Java thread for each
% active object; thus, indirectly resolving the thread performance
% issue.

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.5]{figs/p50000vvm}
\caption{Increasing parallelism in \Crisp for Prime Sieve}
\label{fig:prime_sieve_cacoj}
% \end{center}
% \end{figure}
% 
% \begin{figure}[t]
% \begin{center}
  \includegraphics[scale=0.44]{figs/p50000sysmon} 
  \caption{Utilizing both CPUs with Prime Sieve in \Crisp}
  \label{fig:multi-cpu-1}
\end{center}
\end{figure}

\section{Related Work} \label{sec:relwork}
\input{related2}

\section{Conclusion} \label{sec:conclusion}
In this paper, we proposed \Crisp as an implementation scheme for
application-level scheduling of active objects. 
\Crisp first introduces asynchronous message passing with fine-grain
priority management and scheduling of messages. 
Additionally, it introduces a Creol to Java compiler that translates
the active objects in Creol into an equivalent Java application. 
\Crisp compiler seamlessly integrates Java class libraries into Creol
including data types that turns Creol from a modeling language to a
fully-fledged one in the hands of the programmer. 
% \Crisp compiler takes advantage of template engines; i.e. the code
% generation templates can easily be customized in the direction of the
% requirements of the problems at hand.

The \jtt{java.util.\-con\-current} package provides useful API for
concurrent programming. Java futures facilitate modeling
asynchronous message passing. However, for processor release
points, we had to preserve threads (using \jtt{Interruptible\-Process})
to allow continuations which leads to their OS-level context switching
that is costly. Moreover, we were tightly directed to use the
out-of-the-box \jtt{Executor\-Service} which is limitedly extensible. We
had no control over the scheduling mechanisms of the internal queue used
in the service implementations. Thus, we needed to re-implement some of
the concepts. Through prototyping \Crisp, we learned that
there are two major
challenges ahead. Firstly, we need to integrate continuations into Java
using a many-messages-to-one-thread mapping model. Secondly, we need
complete control over scheduling of messages and threads in
\jtt{Executor\-Service}'s internal queue. Table \ref{tbl:eval}
summarizes this discussion.

\begin{table}[h]\small
\begin{center}
\begin{tabular}{p{1.5cm} p{2cm} p{2cm} p{1.5cm}}
 & Asynchronous Communication & Processor Release Point & Scheduling \\
\thickhline
Modeling & \ding{51} & \ding{51} & \ding{53} \\
Performance & \ding{51} & \ding{53} & \ding{51} \\ \thickhline
 & Java Futures & Interruptible Process & Executor Service \\
\end{tabular}
\caption{Overview of evaluation of challenges}
\label{tbl:eval}
\end{center}
\end{table}

In future, we will focus on thread performance
for \Crisp such that thread scalability can be achieved to a certain
limit. Additionally, the development of concurrency features on
multi-core in \Crisp is one of the major future concentrations.
% We will extend Creol to provide the necessary syntax to express all
% different levels of priorities.
% including method level, message level, and object level priorities. 
Moreover, another line of future work involves profiling and monitoring
objects  at runtime  to be used for optimization and performance
improvement.
In addition, we intend to extend and integrate into our tool set 
model checking engines such as Modere \cite{JaghooriMS06}.

% \bibliographystyle{plain}
% \bibliography{references}

% Local Variables:
% fill-column: 80
% End: