### 2015-10-20 p01-ch01-2012/problem.tex

In this section we briefly sketch the main software engineering problem of deploying 
application-level scheduling of Creol objects onto multicore.

A simple implementation  maps each Creol object unto a single Java thread (e.g., {\jtt{java.lang.Thread}}) which
executes messages taken from a message queue according to some
scheduling policy.  The major drawback of this implementation strategy
is scalability.  OS-level threads are costly resources, and a typical
system written in object-oriented style might instantiate tens of
thousands of objects at runtime.  Thus, any real-world program would
quickly overwhelm current operating systems, see for example our case study in Section \ref{sec:caseStudy}.
Another major software engineering problem of this scheme is that it requires an implementation of  asynchronous method calls 
(e.g., polling/waiting for return values and storing/retrieving local variables).

The basic mechanism of asynchronous method calls is in fact provided by the {\jtt{java.util.\-concurrent}} package by
means of thread pools
which leverage hardware level constructs  to allow a reduction of the overhead of the multi-threaded Java programs by using lock-free, and wait-free thread-safety mechanisms.
This package, e.g., the implementation of {\jtt{FutureTask}}, requires another implementation strategy for a Java backend for Creol which uses
one Java thread per Creol process and allows for a one-to-one translation of Creol classes and methods to Java classes and methods.
However,  one of the major problems in leveraging scheduling issues is that this package does not allow high-level control of the scheduling of the threads.
The main contribution of the paper is the introduction of a general architecture on top of the {\jtt{java.util.\-concurrent}} package
to provide such high-level control for concurrent objects.

% Local Variables:
% mode: LaTeX
% TeX-master: "paper.tex"
% End: