### 2015-10-20 p03-ch01-2015-monitoring/paper-esocc.tex

%!TEX root = ../thesis.tex

\chapter[Formal verification of service level agreements through distributed monitoring]{Formal verification of service level agreements through distributed monitoring%
\footnote{This paper is funded by the EU project FP7-610582 ENVISAGE: Engineering Virtualized
Services, \url{http://www.envisage-project.eu}.}}
% 
\label{ch:p03:ch01}
% 
\chapterauthor{Behrooz Nobakht, Stijn de Gouw, Frank S. de Boer}

\section*{Abstract}
\label{abstract}
In this paper, we introduce a  formal model of the  availability, budget compliance
and  sustainability of distributed services, where service   sustainability is
a new concept which arises as 
% an extension of 
the composition of 
service  availability and budget compliance.
The model formalizes a distributed  platform for monitoring the above
service characteristics in terms of a parallel composition of task automata,
where dynamically generated tasks model asynchronous events with deadlines.
The main result of this paper is a formal model to optimize and reason about service characteristics
through monitoring.
In particular, we use schedulability analysis of the
underlying timed automata to optimize and guarantee service sustainability.

%to capture characteristics of a distributed deployment environment.
%In the distributed deployment, a service runs across several resources in a virtualized environment.
%We formalize characteristics of a distributed service by service availability and service budget compliance.
%We introduce service sustainability as an extension and composition of the two.
%We design timed automata to ensure the given service level agreement for each characteristic.
%The formal model is realized by devising an actor-based distributed monitoring platform using the Abstract Behavioral Specification language ABS.
%We discuss how the characteristics of the model can be utilized in the application and business in different aspects.
%In addition, we discuss how monitoring model parameters; e.g. monitoring window; influence maintaining service characteristics in a distributed environment.
% We present algorithms to generate the monitoring platform components given input parameters in order to guarantee the defined service level agreements.

% \keywords{runtime monitoring, service availability, budget compliance, service sustainability, distributed architecture, cloud computing, service level agreement}

\input{sec01-intro}
\input{sec02-related}
\input{sec03-frh-example}
\input{sec045-defs}
\input{sec05-monitoring-model}
\input{sec06-impl}
\input{sec07-conclusion}

% \bibliographystyle{abbrv}
% \bibliography{refs}

% \end{document}