### 2015-11-24 p03-ch02-2014-jmseq/grammar-sample-usage-1.tex
Add JMSeq Journal Paper from SCP 2014

\documentclass{article}
\usepackage[top=0.01in,bottom=0.01in,left=0.01in,right=0.01in,paperwidth=5.1in,paperheight=0.4in]{geometry}

\usepackage{alltt}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
%\usepackage{courier}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{url}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

\begin{document} 
\pagestyle{empty}
\begin{alltt} 
\centering\small
{\color{red}\{}call(* *.A.m_a(..)){\color{red}[}{\color{blue}\{}call(* *.B.m_b(..)){\color{blue}[}\{call(* *.C.m_c(..))\}{\color{blue}]}{\color{blue}\}}{\color{red}]}{\color{red}\}}
\end{alltt}
\end{document}