### 2015-11-24 p03-ch02-2014-jmseq/grammar-sample-usage-3.tex

\documentclass{article}
\usepackage[top=0.01in,bottom=0.01in,left=0.01in,right=0.01in,paperwidth=5.1in,paperheight=0.4in]{geometry}

\usepackage{alltt}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
%\usepackage{courier}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{url}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

\begin{document} 
\pagestyle{empty}
\begin{alltt} 
\centering\small
\{call(* *.A.m_a(..))\}\${\color{red}<}\{call(* *.B.m_b(..))\}\#{\color{red}?}\{call(* *.C.m_c(..))\}\#{\color{red}>}
\end{alltt}
\end{document}