### 2016-02-08 cv.tex
Add single-page CV to the thesis
While there, re-order some of the sections based on Leiden U ordering.

% -*- program: lualatex -*-

\documentclass[a4paper,11pt]{moderncv}
\moderncvtheme[grey]{classic}

% === Packages
\usepackage[scale=0.8]{geometry}
\usepackage[utf8]{inputenc}
% \usepackage{hyperref}              
% \definecolor{linkcolour}{rgb}{0,0.2,0.6}
% \hypersetup{colorlinks,breaklinks,urlcolor=linkcolour, linkcolor=linkcolour}

% === Fonts
\usepackage[T1]{fontenc}
\usepackage{charter}

% === Bullet Style
\def\labelitemi{{\tiny$\bullet$}}

% === CV Information
\title{ }
\firstname{Curriculum Vitae}
\familyname{ }
\address{ }{ }{ }
% \phone[mobile]{+31 611 770 610}
\email{behrooz.nobakht@gmail.com}
% \social[twitter]{behruz}
% \social[linkedin]{behrooznobakht}
\extrainfo{Geboren 29-07-1981, Teheran}
% \homepage{LinkedIn}

\begin{document}
\maketitle

% Education
\section{Opleiding}
%\cventry{year--year}{Degree}{Institution}{City}{\textit{Grade}}{Description}

\cventry{2011--2015}{PhD Informatica}{Universiteit Leiden}{Leiden}{}
{\emph{Actors At Work}\copyright}

\cventry{2009--2011}{MSc cum laude Informatica}{Universiteit Leiden}{Leiden}{}{\emph{Deployment of active objects onto multi-core}}

\cventry{1998--2002}{BSc Informatica}{Technologie Universiteit Sharif}{Teheran}{\textit{7.0}}{\emph{Online Shopping}}

% Work Experience
\section{Werkervaring}
% \cventry{year--year}{Job title}{Employer}{City}{}{General description no%
% longer than 1--2 lines.\newline{}%
% Detailed achievements:%
% }

\cventry{2011--Nu}{Senior Software Ontwikkelaar}{SDL Fredhopper}{Amsterdam}{}{Groep leider in cloud\newline}

\cventry{2003--2009}{Medeoprichter en architect}{ASTA Co.}{Teheran}{}{%
\begin{itemize}
\item Het beheer van het bedrijf
\item Ontwikkelen
\item Project onderhandeling
\end{itemize}
}

\end{document}