### 2016-05-04 abstract.tex
abstract: removed duplicate word

% !TEX root = thesis.tex
%
\pdfbookmark[0]{Abstract}{Abstract}
\chapter*{Abstract}
\label{sec:abstract}
\vspace*{-10mm}

This thesis contributes to the intersection of object orientation, actor model, and concurrency.
We choose Java as the main programming language and as one of the mainstream 
object-oriented languages. 
We formalize a subset of Java and its concurrency API to 
facilitate formal verification and reasoning about it.
We create an abstract mapping from a concurrent-object modeling language, 
ABS, to the programming semantics of concurrent Java. 
We provide the formal semantics of the mapping and runtime properties of 
the concurrency layer including deadlines and scheduling policies.
We provide an implementation of the ABS concurrency layer as a Java API library 
and framework utilizing the latest language additions 
in Java~8.
We design and implement a runtime monitoring framework, JMSeq, to verify the
correct ordering of execution of methods through code annotations in JVM. 
In addition, we design a large-scale monitoring system as a real-world 
application; the monitoring system is built with ABS concurrent objects 
and formal semantics that leverages schedulability 
analysis to verify correctness of the monitors.

\vspace*{20mm}

{\usekomafont{chapter}Samenvatting}\label{sec:abstract-diff} \\

\TODO{Translate to Dutch. Kindly by Stijn.}