### 2016-05-04 p03-ch01-2015-monitoring/sec55-app.tex

% -*- root: paper.tex -*-

\section{Applications and Future Work} % (fold)
\label{ch04:sec:application}

\subsection{Code Generation for Monitors}

The parameters $E(s,\tau)$ and $B(s,\tau)$ are significantly driven from a business perspective.
Such parameters enforce business-level constraints on the services that are also bound from a legal perspective.
In the past decade, there is considerable research on how to specify and model such business constraints generally referred to as service level agreements (SLA).
A number of languages, frameworks and models have been developed to be able to model and specify SLAs for the deployment of services.
Among others, we refer to SLA$\star$~\cite{kearney2010sla}, WSLA~\cite{keller2003wsla}, and SLAng~\cite{lamanna2003slang}.
The focus of this research is not to enter this domain and yet build on top of it a model to be able to reason about ``distributed service properties'' such as service availability or budget compliance. 
In the following, we present a set of pseudo algorithms for the model in Section \ref{ch04:sec:modelling} that present how the monitoring can be generated based on the agreements that are specified
\footnote{This is a work in progress in ENVISAGE project WP2}.

In Algorithm \ref{ch04:alg:gen:monitor}, the monitoring actor method \jtt{check} is presented that collects the measurements and publishes them to the platform.

% 
\begin{algorithm}
\caption{Monitor for service $s$}
\label{ch04:alg:gen:monitor}
\begin{algorithmic}
\Function{check}{}  
  \For{$r \in \sigma(s)$}
    \State $p \; ! \; \jtt{publish}(\mu(s,r,\tau))$
    \State $\bfjtt{this} \; ! \; \jtt{check}() \; \bfjtt{duration}(\tau)$
  \EndFor
\EndFunction
\end{algorithmic}
\end{algorithm}
%

When platform receives a measurement, it updates its state, computes the service properties and initiates an asynchronous message to process and react to them.
This is presented in Algorithm \ref{ch04:alg:publish}.

% 
\begin{algorithm}
\caption{Process measurements}
\label{ch04:alg:publish}
\begin{algorithmic}
\Function{publish}{$\mu$}
  \State $\alpha(s,\tau,t_c) \gets$ \Call{compute$_{\alpha}$}{}
  \State $\beta(s,\tau) \gets$ \Call{compute$_{\beta}$}{}
  \State Update $\sigma(s)$ and $h(s)$
  \State $\bfjtt{this} \; ! \;$ \Call{process}{$\mu,\alpha(s,\tau,t_c),\beta(s,\tau)$}
\EndFunction
\end{algorithmic}
\end{algorithm}
%

As a consequence of publishing measurements, the platform react to them in case they do not match the expectations.
The behavior is discussed in Theorem \ref{ch04:thm:sustain:conv}.
Algorithm \ref{ch04:alg:process} presents the pseudo code for the behavior that makes changes to $M_\alpha$ and $M_\beta$.

% 
\begin{algorithm}
\caption{$M_\alpha$ and $M_\beta$}
\label{ch04:alg:process}
\begin{algorithmic}
\Function{process}{$\mu,\alpha(s,\tau,t_c),\beta(s,\tau)$}
  \If{$\alpha(s,\tau,t_c) < 1 - \varepsilon_\alpha$}
    \State $n_r \gets$ \Call{compute$_{n_r}$}{}
    \State $r \gets p \; ! \; \jtt{allocate}(s, n_r)$
  \EndIf
  \If{$\beta(s,\tau) - B(s, \tau) < 1 - \varepsilon_\beta$}
    \State $p \; ! \; \jtt{deallocate}(s)$
  \EndIf
  \State Update $\sigma(s)$ and $h(s)$
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Monitoring Window $\tau$}

Every monitoring measurement is performed in a monitoring window $\tau$.
Monitoring measurements are aggregated and calculated in every window and form the basis of reactions and corrections that are necessary to balance the running services to meet their expectations.
The selection of an appropriate monitoring window length $\tau$ is crucial to manage the deployments of the services according to their agreements. 
The authors in \cite{hogben2013defavail} present that for the same setup and deployment of services, measurements using different monitoring windows yield to very different understanding of service properties such as service availability.
Therefore, it is essential to choose the value of $\tau$ such that monitoring measurements do not lead to unrealistic understanding and inappropriate reactions.
To this end, the following should be considered when selecting a monitoring window $\tau$:
\begin{itemize}
\item We defined $\hat{t}$ to be the time a resources needs to initialized in the platform.
Under Assumption \ref{ch04:def:resource:init:time}, all resources require same amount of time to initialize.
If $\tau < \hat{t}$, Theorem \ref{ch04:def:service:availability} invalidates because the measurements are done in a time window that is not enough to be able to take the newly allocated resource into account for a service.
Thus, it is essential that $\tau \geq \hat{t}$.
It can be similarly argued that choosing monitoring window as $\tau \gg 2 \times \hat{t}$ also has a counter-productive effect on the service deployments.
Widening the monitoring window causes to lose fine granularity of monitoring measurements to be able to react to them in proper time.
\item For simplicity of reasoning, we considered Assumption \ref{ch04:def:single:resource:type};
i.e. all the resources provisioned by the platform are of the same type and as such $\hat{t}$ is the same.
However, different services may fit into different types or resources and using different types of resource may fit better for budget compliance as they cost differently.
In such setting, the monitoring window should be at least the maximum $\hat{t}$ of any resource type that is available in the platform: $\tau \geq \mathsf{max}(\hat{t}) \; \forall r \in p$; otherwise, Theorem \ref{ch04:thm:service:budget:conv} violates.
\end{itemize}

% section applications_and_discussion (end)