### 2016-08-18 acknowledgement.tex
Switch to 'biber' backend for biblio
Fix multiple issues in the bibliography file
and special characters for the author names
from different locales.

Biber supports using 'IDS' to alias multiple
keys for the same citation.

Signed-off-by: Behrooz Nobakht <nobeh5@gmail.com>

%
\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}

There will be acknowledgments.