### 2016-10-04 acknowledgement.tex
Update non-scientific parts
Trying to get ready for publication

%
\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}

To mom, who sacrificed all she could for us and left us very early.
To dad, who never loses hope in us.
To Shima, whose love and patience are my only sources of strength.
To Maryam and Ali, together with who, we never stop making our parents proud.

This thesis like many others has been a journey.
A journey that passed through the lives of family, friends, and colleagues.
This is to extend a warm thankful message to all; especially
Hossein Rahmani, Ramin Etemadi, Mahdi Jaghoori, Mohsen Davarinejad,
Vlad Serbanescu, Nikolaos Bezirg, Keyvan Azadbakht.
And, thank you Mohsen Davarinejad and Hossein Rezazadeh Mehrizi to honor 
me for the defense session.
It was a pleasure for me to meet and work with great minds in the context of
ENVISAGE and HATS projects.

I would like to thank Dr. Peter Wong who accepted to honor me for the opposition
committee of the defense session.