### 2016-10-13 acknowledgement.tex
Update Acknowledgements

%
\pdfbookmark[0]{Acknowledgements}{Acknowledgements}
\chapter*{Acknowledgements}
\label{sec:acknowledgement}
\vspace*{-10mm}

To mom, who sacrificed all she could for us and left us very early.
To dad, who never loses hope in us.
To Shima, whose love and patience are my only sources of strength.
To Maryam and Ali, who selflessly made room for me to get to where I am.

This thesis like many others has been a journey.
A journey that passed through the lives of family, friends, and colleagues.
This is to extend a warm thankful message to all; especially
Hossein Rahmani, Ramin Etemadi, Mahdi Jaghoori, Mohsen Davarinejad,
Vlad Serbanescu, Nikolaos Bezirgiannis, Keyvan Azadbakht.
It was a pleasure for me to meet and work with great minds during
ENVISAGE and HATS projects together with Fredhopper team.
I would like to thank Amir Hajari who has been like a wiser brother 
for me during the years that I have lived in The Netherlands.

I would like to thank Dr. Peter Wong who accepted to honor me for the opposition
committee of the defense session.