### 2016-11-01 acknowledgement.tex
Update Acknowledgements

%
\pdfbookmark[0]{Acknowledgements}{Acknowledgements}
\chapter*{Acknowledgements}
\label{sec:acknowledgement}
\vspace*{-10mm}

To mom, who sacrificed all she could for us and left us very early.
To dad, who never loses hope in us.
To Shima, whose love and patience are my only sources of strength.
To Maryam and Ali, who selflessly made room for me to get to where I am.
To Fatemeh, whose birth has given us passion for life more than we have ever thought.

This thesis like many others has been a journey.
A journey that passed through the lives of family, friends, and colleagues.
This is to extend a warm thankful message to my friends; especially
Hossein Rahmani, Ramin Etemadi, Mohammad Rezazade Mehrizi, Lucas Sakizloglou,
and Mohammad Mahdi Jaghoori.
I would like to thank Amir Hajari who is like a wiser brother 
for me during the years that I have lived in The Netherlands.
It was a pleasure for me to meet and work with great minds during
ENVISAGE and HATS projects together with Fredhopper team.